# Bootstrap {-}

***

```{definition}
Bootstrap strategies compose a class of non-parametric Monte Carlo methods
that estimate the distribution of a population by means of resampling.
```



> Resampling methods treat samples as a finite population, from which
> samples are taken to estimate features and make inferences about this
> population.

To generate a random bootstrap sample by \(x\) resampling, generate \(n\)
evenly distributed random integers in the set \(1, ..., n{}\) and select
the bootstrap sample as \(x^{*} = (x_{i_{1}}, ..., x_{i_{n}})\).

```{example}
Suppose we observe the following data:

\(x = {2, 2, 1, 1, 5, 4, 4, 3, 1, 2}\).

Resampling from \(x\), we select \(\{1, 2, 3, 4, 5\}\) with
probabilities \(\{0.3, 0.3, 0.1, 0.2, 0.1\}\). The \(F_{X^{*}}(x)\)
distribution of a sample taken at random is exactly the function
\(F_{n}\):

\[ F_{X^{*}}(x) = F_{n}(x) = \begin{cases}
                              0.0 & x < 1; \\ 
                              0.3 & 1 \leq x < 2; \\ 
                              0.6 & 2 \leq x < 3; \\ 
                              0.7 & 3 \leq x < 4; \\ 
                              0.9 & 4 \leq x < 5; \\
                              1.0 & x \geq 5.
                             \end{cases}
\]
```

## Standard error boostrap estimate {-}

The bootstrap estimate of the standard error of an estimator
\(\hat{\theta}\) is the sample standard error of the bootstrap 
replicates \(\hat{\theta}^{(1)}, ..., \hat{\theta}^{(B)}\):

\[ \text{se}(\hat{\theta}^{*}) =
   \sqrt{\frac{1}{B - 1}
         \sum_{b = 1}^{B} (\hat{\theta}^{(b)} - \bar{\theta}^{*})^{2}},
         \\
   \bar{\theta}^{*} = \frac{1}{B} \sum_{b = 1}^{B} \hat{\theta}^{(b)}.
\]

Where \(B\) is the number of replicates.

```{example, name="Correlation"}
```

(ref:corr) A scatterplot of the `speed` and `dist` variables of the 
           `cars` dataframe.

```{r, fig.cap='(ref:corr)', fig.align='center', message=FALSE, comment=NA}
lattice::xyplot(dist ~ speed, cars, pch = 16, ylab = list(rot = 0))

set.seed(1)
B <- 10e3
boot.corr <- vector('numeric', B)
for (b in 1:B){
  ind <- sample(nrow(cars), replace = TRUE)
  boot.corr[b] <- with(cars[ind, ], cor(dist, speed))
}
(theta.star <- mean(boot.corr))

(se.theta.star <- sd(boot.corr))
```

## Bias bootstrap estimate {-}

The definition of bias is given by:

\[ B(\hat{\theta}) = E(\hat{\theta}) - \theta. \]

Thus, the bias bootsrap estimate is:

\[ \hat{B(\hat{\theta})} = \bar{\hat{\theta}^{*}} - \theta, \]

expression in which \(\bar{\hat{\theta}^{*}} =
\frac{1}{B} \sum_{b = 1}^{B} \hat{\theta}^{(b)}\) and \(\hat{\theta}\)
is the estimate calculated using the original sample.

```{example, name="Correlation"}
```

```{r, comment=NA}
theta.star - with(cars, cor(dist, speed))
```

***
